### 2014-11-05 intro.tex
edits

Many real-world web services --- such as those built and maintained by
Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
logic across multiple \emph{replicas} within and across data
centers. Replication is intended not only to improve application throughput
and reduce user-perceived latency, but also to tolerate partial failures
without compromising overall service availability. Traditionally programmers
have relied on \emph{strong consistency} guarantees such as
linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct
applications.  While strong consistency is an easily stated property, it
masks the reality underlying large-scale distributed systems with respect to
non-uniform latency (i.e., availability) and network
partitions~\cite{Brewer2000,Gilbert2002}. Indeed, modern web services, which aim
to provide an "always on" experience, overwhelmingly favour availability and
partition tolerance over strong consistency. To this end, several \emph{weak
  consistency} models such as eventual consistency, causal consistency,
session guarantees, and timeline consistency have been proposed.

@@

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a shallow extension of Haskell that supports the description
  and validation of replicated data types found in programs intended to
  execute on eventually consistent stores. Contracts are used to specify
  fine-grained application-level consistency properties, and are analyzed to
  assign the most efficient and sound store consistency level to the
  corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct and most efficient isolation level for a transaction.
\item We provide metatheory that certifies the soundness of our contract
	enforcement system, and ensures that an operation is only executed if the
	required conditions on consistency are met.
\item An implementation of \name as a transparent shim layer over
	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
	extensive experimental evaluation over a collection of real-world
	applications, including a Twitter-like micro-blogging site and an eBay-like
	auction site illustrates the practicality of our approach.
\end{itemize}

>>>>

Many real-world web services --- such as those built and maintained by Amazon,
Facebook, Google, Twitter, etc. --- replicate application state and logic
across multiple \emph{replicas} within and across data centers. Replication is
intended not only to improve application throughput and reduce user-perceived
latency, but also to tolerate partial failures without compromising overall
service availability. Traditionally programmers have relied on \emph{strong
consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
serializability~\cite{Serializability} in order to build correct applications.
While strong consistency is an easily stated property, it masks the reality
underlying large-scale distributed systems with respect to non-uniform latency,
availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
modern web services, which aim to provide an "always on" experience,
overwhelmingly favour availability and partition tolerance over strong
consistency. To this end, several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have been proposed.

@@

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a shallow extension of Haskell that supports the
	description and validation of replicated data types found on eventually
	consistent stores. Contracts are used to specify fine-grained
	application-level consistency properties, and are analyzed to assign the most
	efficient and sound store consistency level to the corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct isolation level for a transaction.
\item We provide metatheory that certifies the soundness of our contract
	enforcement system, and ensures that an operation is only executed if the
	required conditions on consistency are met.
\item An implementation of \name as a transparent shim layer over
	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.
	Experimental evaluation over a set of real-world applications, including a
	Twitter-like micro-blogging site and an eBay-like auction site illustrates
	the practicality of our approach.
\end{itemize}

### 2014-11-05 model.tex
edits

\begin{figure}
\centering
\includegraphics[width=0.75\columnwidth]{Figures/SystemModel}
\caption{Quelea system model.}
\label{fig:sysmod}
\end{figure}

@@

Each object is associated with a set of \emph{operations}. The clients interact
with the store by invoking operations on objects. The sequence of operations
invoked by a particular client on the store is called a \emph{session}. The
data store is typically accessed by a large number of clients (and hence
sessions) concurrently. Importantly, the clients are oblivious to which replica
an operation is applied to; the data store may choose to route the operation to
the closest replica for minimizing latency, or route it to a different replica
in order to balance the load in the system. For example, the operations
\emph{foo} and \emph{bar} invoked by the same session on the same object, might
end up being applied to different replicas because replica 1 (to which
\emph{foo} was applied) might be unreachable when the client invokes
\emph{bar}.

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of effects at that
replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
appended to the state of $x$ at replica 1. Thus, every operation is evaluated
over a \emph{snapshot} of the state of the object on which it is invoked. In
this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$ and $w^x_2$.
Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$, written
logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where $\visZ$ is
the visibility relation between effects. Visibility is an irreflexive (an
operation cannot see its own effect) and asymmetric relation, and only relates
effects produced by operations on the same object. Executing a read-only
operation is similar except that no new effects are produced.

The effect added to a particular replica is asynchronously sent to other
replicas. We assume that \emph{eventually} an effect applied at one replica
will be merged into all other replicas. Two effects $w^x_4$ and $w^x_5$ that
arise from the same session are said to be in \emph{session order} (written
logically as $\so{w^x_4}{w^x_5}$). Session order is an irreflexive, transitive
relation. The effects $w^x_4$ and $w^x_5$ arising from operations applied to
the same object $x$ are said to be under the \emph{same object} relation,
written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every effect with
the operation that generated the effect with the help of a relation $\operZ$.
In the current example, $\oper{w^x_4}{foo}$ and $\oper{w^x_5}{bar}$ hold. For
simplicity, we assume all operation names across all object types are distinct.

>>>>

\begin{figure}
\centering
\includegraphics[width=0.75\columnwidth]{Figures/SystemModel}
\caption{\name system model.}
\label{fig:sysmod}
\end{figure}

@@

Each object is associated with a set of \emph{operations}. The clients interact
with the store by invoking operations on objects. The sequence of operations
invoked by a particular client on the store is called a \emph{session}. The
data store is typically accessed by a large number of clients (and hence
sessions) concurrently. Importantly, the clients are oblivious to which replica
an operation is applied to; the data store may choose to route the operation to
any replica in order to minimize latency, balance load, etc. For example, the
operations \emph{foo} and \emph{bar} invoked by the same session on the same
object, might end up being applied to different replicas because replica 1 (to
which \emph{foo} was applied) might be unreachable when the client invokes
\emph{bar}.

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of effects at that
replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
appended to the state of $x$ at replica 1. Thus, every operation is evaluated
over a \emph{snapshot} of the state of the object on which it is invoked. In
this case, the effects $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
$\visZ$ is the visibility relation between effects. Visibility is an
irreflexive and asymmetric relation, and only relates effects produced by
operations on the same object. Executing a read-only operation is similar
except that no new effects are produced.

The effect added to a particular replica is asynchronously sent to other
replicas, and eventually merged into all other replicas. Two effects $w^x_4$
and $w^x_5$ that arise from the same session are said to be in \emph{session
order} (written logically as $\so{w^x_4}{w^x_5}$). Session order is an
irreflexive, transitive relation. The effects $w^x_4$ and $w^x_5$ arising from
operations applied to the same object $x$ are said to be under the \emph{same
object} relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate
every effect with the operation that generated the effect with the help of a
relation $\operZ$. In the current example, $\oper{w^x_4}{foo}$ and
$\oper{w^x_5}{bar}$ hold. For simplicity, we assume all operation names across
all object types are distinct.
